---
output: html_document
---
### Udacity/ Data Analysis With R
### PS1:  Explore One Variable: "Cholesterol in blood in female population"

### About Dataset

Definition and explanations:

1. Indicator Name: TC female (mmol/L), age standardized mean.

2. Definition of the indicator: The mean TC (Total Cholesterol) of the female population, counted in mmol per L; this mean is calculated as if each country has the same age composition as the world population. 

3. Source organization(s):	School of Public Health, Imperial College London

4. Link to source organization:	http://www.imperial.ac.uk/

5. Complete reference:	MRC-HPA Centre for Environment and Health

6. Link to complete reference:	http://www.imperial.ac.uk/medicine/globalmetabolics/

7. Dataset consists data of the female populations of 199 countries (1980 through 2008).

8. Gapminder Data: https://www.gapminder.org/

9. Auxiliary data: https://github.com/lukes/ISO-3166-Countries-with-Regional-Codes/tree/master/all


### Overview of the Data
Using the Cholesterol (fat) in blood in female population Gapminder data set, some exploratory graphics were built to understand the trend of this indicator. 

Loading libraries:
```{r}
library(gridExtra)
library(ggplot2)
```

Loading csv data file:
```{r}
dt <- read.csv('Indicator_TC_female_ASM.csv', header = T,  check.names = F)
```

### Cleaning and Re-shaping Data
Quick view of the statistics:
```{r}
dim(dt)
str(dt)
```

Changing the first column name:
```{r}
names(dt)[1] <- c("Country")
```

Loading auxiliary data set containing country names and the region /continent/ to which each belongs:
```{r}
country_region <- read.csv('all.csv', header = T,  check.names = F, na.strings=c("", "NA"))
```
```{r}
dim(country_region)
str(country_region)
```

Reshaping, filtering, cleaning and changing the column names of the auxilary data frame:
```{r}
country_region <- country_region[, c(1, 6, 7)]
country_region <-subset(country_region, !is.na(country_region$region))
names(country_region) <- c("Country", 'Region', 'Sub-Region')
```

Merging the main and the country/region data frames - taking the intersection of data:
```{r}
dtm <- merge(dt, country_region, by="Country")
str(dtm)
```

Checking how many countries did not have the equivalent and are left after the merging:
```{r}
setdiff(dt$Country, country_region$Country)
```

### Data Exploration

Creating various subsets of regions or sub-regions for the future investigation:
```{r}
region_americas <- subset(dtm, dtm$'Sub-Region' == 'Northern America' )
region_africa <- subset(dtm, dtm$'Region' == 'Africa' )
region_europe <- subset(dtm, dtm$'Region' == 'Europe' )
```

Creating a histogram of the mean of the Total Cholesterol (TC) of the female population for each region: 
```{r}
ggplot(aes(x = dtm$'2008'), data = dtm,color = I('black'))+
  geom_histogram(binwidth = .2, color="#A9A9A9", fill = '#E69F00')+
  scale_x_continuous(breaks = seq(3.5, 6, .5), limits = c(3.5,6)) +
  facet_wrap(~`Region`, ncol=1)+
  xlab('The mean TC (Total Cholesterol) of the female population')+ 
  ylab('Count')+
  labs(title = "2008 Distribution of the TC mean")
```

Investigating the TC mean of data frame using box plots for each region:
```{r}
ggplot(aes(x = dtm$Region, y = dtm$'2008'), data = dtm)+
  geom_boxplot(outlier.color='red', color="#A9A9A9", fill = '#E69F00')+
  xlab('Regions')+
  ylab('The TC mean')+
  labs(title = "2008 Distribution of the TC mean")
```


```{r}
region_africa_2008 <- region_africa[order(-region_africa$'2008'),]
qplot(region_africa_2008$'2008',region_africa[1:17]$Country)+
  xlab('The mean TC (Total Cholesterol) of the female population')+ 
  ylab('Countries')+
  labs(title = "Country TC mean (Africa region)")
```


### Resources
http://www.ats.ucla.edu/stat/r/faq/subset_R.htm

http://www.statmethods.net/

http://flowingdata.com/2015/02/18/loading-data-and-basic-formatting-in-r/

http://www.statmethods.net/input/datatypes.html

http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/

http://docs.ggplot2.org/current/scale_continuous.html

http://flowingdata.com/2014/02/27/how-to-read-histograms-and-use-them-in-r/


