---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Udacity / Data Analysis With R
### PS1:  Exploring Your Friends' Birthdays
****
### Assignment
The task is to investigate the distribution of my friends' birth months and days and answer questions:

1. How many people share your birthday?

2. Which month contains the most number of birthdays?

3. How many birthdays are in each month?

4. Which day of the year has the most number of birthdays?

5. Do you have at least 365 friends that have birthdays on every day of the year?

### About Dataset
Working with a list of birthdays in provided birthdays example file. 
***
### Overview of the Data
Loading libraries:
```{r}
library(ggplot2)
library(lubridate)
```
***
Loading the data file:
```{r}
dt <- read.csv('birthdaysExample.csv')
```
***
Quick view of the statistics:
```{r}
dim(dt)
str(dt$dates)
head(dt)
names(dt)
```
***
Importing Dates from Character Format:
```{r}
betterDates <- as.Date(dt$dates, format = "%m/%d/%y")
head(betterDates)
```
***
Statistics of the formatted data:
```{r}
mean_date <- mean(betterDates)
max_date <- max(betterDates)
min_date <- min(betterDates)
mean_date
max_date
min_date
typeof(betterDates)
```
***
#### 1. How many people share your birthday?
```{r}
bd_shared <- dt[betterDates == "2014-12-07",]
bd_shared
length(bd_shared)
```
***
#### 2. Which month contains the most number of birthdays?
```{r}
bd_months <- months(betterDates)
max(table(bd_months))
bd_month_df <- as.data.frame(table(bd_months))
max_month_row = bd_month_df$Freq == max(bd_month_df$Freq)
bd_month_df[max_month_row,]
```
***
#### 3. How many birthdays are in each month?
```{r}
table(bd_months)
```
***
#### 4. Which day of the year has the most number of birthdays?
```{r}
bd_days <- day(betterDates)
bd_df <- as.data.frame(table(bd_days))
max_day_row = bd_df$Freq == max(bd_df$Freq)
bd_df[max_day_row,]
```
***
#### 5. Do you have at least 365 friends that have birthdays on every day of the year?
```{r}
dim(unique(dt[1]))[1] >= 365
```
***
### Resources
https://www.r-bloggers.com/date-formats-in-r/

https://www.gcal2excel.com/

https://support.google.com/calendar/answer/37111?hl=en

https://cran.r-project.org/web/packages/lubridate/lubridate.pdf